---
title: "PDX Bike Count Analysis"
output: html_document
---
by Kevin 

This document summarizes the work that was completed during the Introduction to Scientific Computing for Planners, Engineers, and Scientists course offered by 
PSU in August 2018.

###Data Loading
We first needed to specify the packages that we would be working with.

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

Next, we created a function that would load bike count data in Excel format.
Note that the `date` column was mutated in order to simplify later joining with
weather data.

```{r}
load_bridge <- function(bridge_name) {
  bridge_data <- read_excel(
    "../data/Hawthorne Tilikum Steel daily bike counts 073118.xlsx",
    sheet = bridge_name, skip = 1)
  bridge_data <- add_column(bridge_data, name = bridge_name) %>%
  mutate(date = as.Date(date, format = '%m/%d/%Y'))
  return(bridge_data)
}

bridges <- c("Hawthorne", "Tilikum", "Steel")
all_bridges <-  lapply(bridges, load_bridge)
Hawthorne <- all_bridges[[1]] %>% 
  rename(westbound = `north side (westbound)`) %>%
  rename(eastbound = `south side (eastbound)`) 

Tilikum <-  all_bridges[2]

Steel <- all_bridges[[3]] %>%
  select(-c(`River Walk (lower 2-way)`)) %>% 
  rename(westbound = `north side (westbound)`) %>%
  rename(eastbound = `south side (eastbound)`) 
```
Next, we needed to join bridge and weather data along the date columns.
```{r, message=FALSE}
weather_data_csv <- read_csv("../data/NCDC-CDO-USC00356750.csv") %>% 
  select(-c(STATION, NAME))
  
consolidated_bridge <- bind_rows(Hawthorne, Tilikum, Steel)

all_data <- left_join(consolidated_bridge, weather_data_csv, by = c("date" = "DATE"))
```

### Visualizing data with ggplot2
We were able to visualize data using the recommended ggplot package.
```{r}
all_data %>%
 ggplot(aes(x=PRCP, y=total, color=name)) + 
 theme_minimal() +
 geom_point(size=.5) +
 stat_smooth(method=lm, se=FALSE, linetype="dashed", aes(group=1), 
             size=1, color="black") +
 labs(title= "Total Bike Counts by Precipitation") +
 scale_color_brewer(palette="Dark2")
```
Also fascinating is the visualized effect of the Tilikum crossing based when
juxtaposed against the Hawthorne bridge.

```{r}
tilikum_open <- all_data  %>%
  filter(name == 'Hawthorne' | name == 'Tilikum')  

tilikum_open %>%  ggplot(aes(x=date, y=total, color=name)) + 
  theme_minimal() +
  geom_point(size=.5) +
  stat_smooth(data=subset(all_data, name == "Hawthorne"), method=lm, se=FALSE, linetype="dashed", aes(group=1), 
              size=1, color="black") +
  labs(title= "Total Bike Counts over Time") +
  scale_color_brewer(palette="Dark2")

bike_series <- all_data %>%
  select(date, total, name) %>% 
  group_by(date) %>% 
  summarize(total = sum(total))
```

#### Sub-subheading
Blah